A Hitchhiker's Guide to Reproducible Research in R

### 2021-07-20 manuscript.Rmd
address some comments from #165

**Version control** prevents the ambiguity that arises when multiple versions of code and data are created in parallel during the development phase of a research project.
Version control allows a clear link between which results were generated by which version of code and data.
We recommend using **Git** for version control, because of its widespread adoption in the R community.
Git tracks changes to all project-related files (e.g., materials, data, and code) over time.
At any stage, individual files or the entire project can be compared to, or reverted to, an earlier version.
A version-controlled project makes the loss of files unlikely and also supports remote collaboration.
Git is built around snapshots that represent the project state at a given point in time.
Those snapshots are called "commits" and work like a "save" action.
Changes to files are collected in a _commit_.
Ideally, each commit has a message that succinctly describes these changes.
It is good practice to make commits for concrete milestones, e.g., "Commented on Introduction", "Added SES as a covariate", "Address Reviewer 2's comment 3".
This makes it easier to revert specific changes than when multiple milestones are joined in one commit, e.g., "Changes made on 19/07/2021".
Each commit "knows" its ancestor, and they form a timeline together.
The entirety of commits (i.e. the version-controlled project) is called a repository.
In Git, specific snapshots of a repository can be tagged, such that one can clearly label which version of the project was used to create a preregistration, preprint, or final version of the manuscript as accepted by a journal.
Git has additional features beyond basic version control, such as branches to facilitate collaboration.
@vuorreCuratingResearchAssets2018 provide a more extensive treatment how Git functions and how to use Git for research.
@bryanExcuseMeYou2018 provides additional information on how to track R Markdown documents.
Collaborating via Git requires the repository to be uploaded somewhere.
We recommend GitHub to host Git repositories because of its popularity among R users.
GitHub has many tools that ease project management and collaboration, and these tools provide much value in our everyday work, but they are not central to achieving reproducibility.

@@

A version controlled dynamic document with dependency tracking may still rely on external software.
Troubleshooting issues specific to a particular programming language or dependent tool typically requires considerable expertise and pose a threat to reproducibility.
One solution to avoid problems caused by missing or inconsistent external software dependencies is to provide not only the analysis script but all dependent software packages and system libraries via **software management**.
One comprehensive approach to software management is containerization.
"packaging the key elements of the computational  environment  needed  to  run  the  desired  software [makes] the software much easier to use, and the results easier to reproduce" [@silverSoftwareSimplified2017, p. 174]. <!--APA requires no line number, but pagenumber or section heading-->

**Docker** is an example of containerization. It provides all software used in an analysis, including the operating system itself.
The operating system level is important because some functionality may be passed to software layers beneath the programming language, such as calls to random number generators, linear algebra libraries and such.
Docker does this without interfering with the already installed software by using a virtual software environment independent of the host software environment.
Researchers therefore have the opportunity to use the exact same software setup that was used by the original authors.
Such a snapshot of the software stack is called an "image".
Packaging all required software in such an image requires considerable amounts of storage space.
Two major strategies help to keep the storage requirements reasonable.
One strategy is to rely on pre-made images, which are maintained by a community for particular purposes.
For example, there are pre-made images that only include what is necessary for R, based on Ubuntu (a Linux operating system) containers [@boettigerIntroductionRockerDocker2017].
Users can then install whatever they need in addition to what is provided by these pre-compiled images.
`r ifelse(is.na(container_size), "", glue::glue("The image that was used for this article uses {round(container_size, 2)}Gb of disk space."))`
Our image <!--CJ What is our image? I don't think you introduced this-->includes Ubuntu, R, R Studio, LaTeX as well as a variety of R packages like the tidyverse <!--CJ Yuck-->[@tidyverse] and all their dependent packages, amounting to `r installed.packages() %>% as.data.frame() %>% filter(is.na(Priority)) %>% nrow()` R packages.
A second strategy is to save a container recipe; a so-called  `Dockerfile` containing a textual description of all commands that need to be executed to recreate the software environment, instead of the software environment itself.
Dockerfiles are tiny (the [Dockerfile](https://github.com/aaronpeikert/repro-tutorial/blob/main/Dockerfile) for this project has a size of only `r round(as.numeric(fs::file_size("Dockerfile"))/1024^1, 2)`Kb).
However, they rely on the assumption that all software repositories that provide the dependent operating systems, pieces of software, and R packages will remain accessible in future and will continue to make available historic software versions.
For proper archival, we therefore recommend to store a complete image of the software environment, in addition to the Dockerfile.
A more comprehensive overview of the use of containarization in research projects is given by @wiebelsLeveragingContainersReproducible2021.

>>>>

**Version control** prevents the ambiguity that arises when multiple versions of code and data are created in parallel during the development phase of a research project.
Version control allows a clear link between which results were generated by which version of code and data.
We recommend using **Git** for version control, because of its widespread adoption in the R community.

Git tracks changes to all project-related files (e.g., materials, data, and code) over time.
At any stage, individual files or the entire project can be compared to, or reverted to, an earlier version.
A version-controlled project makes the loss of files unlikely and also supports remote collaboration.
Git is built around snapshots that represent the project state at a given point in time.
Those snapshots are called "commits" and work like a "save" action.
Changes to files are collected in a _commit_.
Ideally, each commit has a message that succinctly describes these changes.
It is good practice to make commits for concrete milestones, e.g., "Commented on Introduction", "Added SES as a covariate", "Address Reviewer 2's comment 3".
This makes it easier to revert specific changes than when multiple milestones are joined in one commit, e.g., "Changes made on 19/07/2021".
Each commit "knows" its ancestor, and they form a timeline together.
The entirety of commits (i.e. the version-controlled project) is called a repository.
In Git, specific snapshots of a repository can be tagged, such that one can clearly label which version of the project was used to create a preregistration, preprint, or final version of the manuscript as accepted by a journal.
Git has additional features beyond basic version control, such as branches to facilitate collaboration.
@vuorreCuratingResearchAssets2018 provide a more extensive treatment how Git functions and how to use Git for research.
@bryanExcuseMeYou2018 provides additional information on how to track R Markdown documents.
Collaborating via Git requires the repository to be uploaded somewhere.
We recommend GitHub to host Git repositories because of its popularity among R users.
GitHub has many tools that ease project management and collaboration, and these tools provide much value in our everyday work, but they are not central to achieving reproducibility.

@@

A version controlled dynamic document with dependency tracking may still rely on external software.
Troubleshooting issues specific to a particular programming language or dependent tool typically requires considerable expertise and pose a threat to reproducibility.
One solution to avoid problems caused by missing or inconsistent external software dependencies is to provide not only the analysis script but all dependent software packages and system libraries via **software management**.
One comprehensive approach to software management is containerization.
"packaging the key elements of the computational  environment  needed  to  run  the  desired  software [makes] the software much easier to use, and the results easier to reproduce" [@silverSoftwareSimplified2017, p. 174].

**Docker** is an example of containerization.
It provides all software used in an analysis, including the operating system itself.
The operating system level is important because some functionality may be passed to software layers beneath the programming language, such as calls to random number generators, linear algebra libraries and such.
Docker does this without interfering with the already installed software by using a virtual software environment independent of the host software environment.
Researchers therefore have the opportunity to use the exact same software setup that was used by the original authors.
Such a snapshot of the software stack is called an "image".
Packaging all required software in such an image requires considerable amounts of storage space.
Two major strategies help to keep the storage requirements reasonable.
One strategy is to rely on pre-made images, which are maintained by a community for particular purposes.
For example, there are pre-made images that only include what is necessary for R, based on Ubuntu (a Linux operating system) containers [@boettigerIntroductionRockerDocker2017].
Users can then install whatever they need in addition to what is provided by these pre-compiled images.
`r ifelse(is.na(container_size), "", glue::glue("The image that was used for this article uses {round(container_size, 2)}Gb of disk space."))`
The image for this project includes Ubuntu, R, R Studio, LaTeX as well as a variety of R packages like the tidyverse <!--CJ Yuck-->[@tidyverse] and all their dependent packages, amounting to `r installed.packages() %>% as.data.frame() %>% filter(is.na(Priority)) %>% nrow()` R packages.
A second strategy is to save a container recipe; a so-called  `Dockerfile` containing a textual description of all commands that need to be executed to recreate the software environment, instead of the software environment itself.
Dockerfiles are tiny (the [Dockerfile](https://github.com/aaronpeikert/repro-tutorial/blob/main/Dockerfile) for this project has a size of only `r round(as.numeric(fs::file_size("Dockerfile"))/1024^1, 2)`Kb).
However, they rely on the assumption that all software repositories that provide the dependent operating systems, pieces of software, and R packages will remain accessible in future and will continue to make available historic software versions.
For proper archival, we therefore recommend to store a complete image of the software environment, in addition to the Dockerfile.
A more comprehensive overview of the use of containarization in research projects is given by @wiebelsLeveragingContainersReproducible2021.
